---
title: The t-distribution and Association Tests
title_meta: Chapter 8
description: In this chapter, you will learn about the t-distribution, association tests, and the chi-squared test.
---
## Exercise 1

We know that with a normal distribution only 5% of values are more than 2 standard deviations away from the mean. Use the `pt` function to calculate the probability of seeing t-distributed random variables being more than 2 in absolute value when the degrees of freedom are 3.

```{r}
1 - pt(2, 3) + pt(-2, 3)
```

## Exercise 2
Now use `sapply` to compute the above probability for degrees of freedom from 3 to 50. Make a plot and notice when this probability converges to the normal distribution's 5%.

```{r}
df <- seq(3,50)
probs <- sapply(df, function(x) 1 - pt(2, x) + pt(-2, x))
plot(df, probs)
```

## Exercise 3
In a previous section, we took random samples from the height distribution of size 50 and noticed that confidence intervals were almost 95%.

```{r}
library(dslabs)
data(heights)
x <- heights %>% filter(sex == "Male") %>%
  .$height
set.seed(1)
mu <- mean(x)
set.seed(1)
N <- 50
B <- 10000
res <- replicate(B, {
  X <- sample(x, N, replace=TRUE)
  interval <- mean(X) + c(-1,1)*qnorm(0.975)*sd(X)/sqrt(N)
  between(mu, interval[1], interval[2])
})
mean(res)
```

    Re-do the above Monte Carlo simulation, but now instead of $N=50$, use $N=15$. Notice what happens to the proportion of hits.
    

```{r}
set.seed(1)
mu <- mean(x)
set.seed(1)
N <- 15
B <- 10000
res <- replicate(B, {
  X <- sample(x, N, replace=TRUE)
  interval <- mean(X) + c(-1,1)*qnorm(0.975)*sd(X)/sqrt(N)
  between(mu, interval[1], interval[2])
})
mean(res)
```

## Exercise 4
Now repeat exercise 2, but since $N=15$ is not that big and we know that heights are normally distributed the t-distribution should apply, use the t-distribution instead of using the normal to construct the confidence intervals.


```{r}
set.seed(1)
N <- 15
B <- 10000
res <- replicate(B, {
  X <- sample(x, N, replace=TRUE)
  interval <- mean(X) + c(-1,1)*qt(0.975, N-1)*sd(X)/sqrt(N)
  between(mu, interval[1], interval[2])
})
mean(res)
```

## Exercise 5
Why did the t-distribution confidence intervals work so much better?

    A. Because the t-distribution takes the variability introduced by estimating the standard deviation into account and generaates larger confidence intervels.
    B. Because the t-distribution shifts the intervals in the right direction.
    C. This was just a chance occurrence. If we run it again, the CLT will work better.
    D. The t-distribution is alwasy a better approximation than the normal.
    
    
    
