---
title: Parameters and Estimates
title_meta: Chapter 2
description: In this chapter, you will learn about parameters and estimates using the example of election polling.
---
## Exercise 1

```yaml
type: MultipleChoiceExercise
lang: r
xp: 100
skills: 1
```

Suppose you poll a population in which a proportion $p$ of voters are Democrats and $1-p$ are Republicans. Your sample size is $N = 25$. Consider the random variable $S$, which is the **total** number of Democrats in your sample. 

What is the expected value of this random variable $S$?

`@instructions`
- $$\mbox{E}(S) = 25(1-p)$$
- $$\mbox{E}(S) = 25p$$
- $$\mbox{E}(S) = \sqrt{25 p (1-p)}$$
- $$\mbox{E}(S) = p$$

`@hint`
- $$\mbox{E}(S)$$ is a function of a function of $p$ and the sample size.

`@sct`
```{r}
msg1 = "Try again. This selection tells you the expected value of a random variable for the number of Republicans in your sample."
msg2 = "Nice work!"
msg3 = "Try again. The expected value of `S` is related to the sample size and the proportion of Democrats in your sample."
msg4 = "Try again. You also need to consider the sample size."
test_mc(correct = 3, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

## Exercise 2

```yaml
type: MultipleChoiceExercise
lang: r
xp: 100
skills: 1
```

Again, consider the random variable $S$, which is the **total** number of Democrats in your sample of 25 voters. The variable $p$ describes the proportion of Democrats in the sample, whereas $1-p$ describes the proportion of Republicans.

What is the standard error of $S$?

`@instructions`
- $$\mbox{SE}(S) = 25 p (1-p)$$
- $$\mbox{SE}(S) = \sqrt{25p}$$
- $$\mbox{SE}(S) = 25(1-p)$$
- $$\mbox{SE}(S) = \sqrt{25 p (1-p)}$$


`@hint`
- The standard error is a function of `p` and the sample size.


`@sct`
```{r}
msg1 = "Try again. The units of variance are squared."
msg2 = "Try  again. The standard error is affected by the proportions of both Republicans and Democrats."
msg3 = "Try again. The standard error is affected by the proportions of Democrats and Republicans."
msg4 = "Nice work!"
test_mc(correct = 4, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

## Exercise 3

```yaml
type: MultipleChoiceExercise
lang: r
xp: 100
skills: 1
```

Consider the random variable $S/N$, which is equivalent to the sample average that we have been denoting as $\bar{X}$. The variable $N$ represents the sample size and $p$ is the proportion of Democrats in the population.

What is the expected value of the $\bar{X}$?

`@instructions`
- $$ \mbox{E}(\bar{X}) = p $$
- $$ \mbox{E}(\bar{X}) = Np $$
- $$ \mbox{E}(\bar{X}) = N(1-p)$$
- $$ \mbox{E}(\bar{X}) = 1-p $$ 

`@hint`
- The expected value of the sum $S$ is $N \times p$. Dividing by the non-random constant $N$ gives us the expected value of the average.

`@sct`
```{r}
msg1 = "Nice work!"
msg2 = "Try  again. The expected value of the sum `S` must be divided by the sample size."
msg3 = "Try again. The expected value is a function of `p`."
msg4 = "Try again. The expected value is a function of `p`."
test_mc(correct = 1, feedback_msgs = c(msg1,msg2,msg3,msg4))
```


## Exercise 4

```yaml
type: MultipleChoiceExercise
lang: r
xp: 100
skills: 1
```

What is the standard error of the sample average, $\bar{X}$? The variable $N$ represents the sample size and $p$ is the proportion of Democrats in the population.

`@instructions`
- $$\mbox{SE}(\bar{X}) = \sqrt{Np (1-p)}$$
- $$\mbox{SE}(\bar{X}) = \sqrt{p (1-p) / N}$$
- $$\mbox{SE}(\bar{X}) = \sqrt{p (1-p)}$$
- $$\mbox{SE}(\bar{X}) = \sqrt{N}$$

`@hint`
- The standard error decreases as the sample size increases.

`@sct`
```{r}
msg1 = "Try again. The standard error of the average gets smaller as you increase the sample size."
msg2 = "Nice work!"
msg3 = "Try again. Remember that you divide by the sample size when calculating the average."
msg4 = "Try again. The expected value is a function of `p`."
test_mc(correct = 2, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

## Exercise 5

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
```

Write a line of code that calculates the standard error `se` of a sample average when you poll 25 people in the population. Generate a sequence of 100 proportions of Democrats `p` that vary from 0 (no Democrats) to 1 (all Democrats).   

Plot `se` versus `p` for the 100 different proportions.

`@instructions`
- Use the `seq` function to generate a vector of 100 values of `p` that range from 0 to 1.
- Use the `sqrt` function to generate a vector of standard errors for all values of `p`.
- Use the `plot` function to generate a plot with `p` on the x-axis and `se` on the y-axis.

`@hint`
- The standard error of the sample average is $$\mbox{SE}(\bar{X}) = \sqrt{p (1-p) / N}$$.

`@pre_exercise_code`
```{r}
#no pec
```

`@sample_code`
```{r}
# `N` represents the number of people polled
N <- 25

# Create a variable `p` that contains 100 proportions ranging from 0 to 1


# Create a variable `se` that contains the standard error of each sample average


# Plot `p` on the x-axis and `se` on the y-axis

```

`@solution`
```{r}
# `N` represents the number of people polled
N <- 25

# Create a variable `p` that contains 100 proportions ranging from 0 to 1
p <- seq(0, 1, length = 100)

# Create a variable `se` that contains the standard error of each sample average
se <- sqrt(p*(1-p)/N)

# Plot `p` on the x-axis and `se` on the y-axis
plot(p, se)
```

`@sct`
```{r}
test_error()

test_function("seq", args = c(from,to,length.out),
              not_called_msg = "Have you used `seq` to generate a vector of 100 values of `p`?",
              args_not_specified = "Remember to specify the starting value, ending value, and length when using `seq` to generate a vector of numbers.",
              incorrect_msg = "Make sure to include the starting value, ending value, and total length of the vector you want to create using the `seq` function.")

test_function("sqrt", args = x,
              not_called_msg = "Have you used `sqrt` to calculate the standard error?",
              incorrect_msg = "Make sure to specify the correct formula when using the `sqrt` function to calculate standard error.")

test_function("plot", args = c(x,y),
              not_called_msg = "Have you used `plot` to generate a plot `p` and `se`?",
              args_not_specified = "Remember to specify the vectors that contain the values to be plotted on the x- and y-axes when using the `plot` function.",
              incorrect_msg = "Make sure to plot `p` on the x-axis and `se` on the y-axis.")

success_msg("Great work! Let's work on a similar problem.")
```


## Exercise 6

```yaml
type: NormalExercise
lang: r
xp: 100
skills: 1
```

Using the same code as in the previous exercise, create a for-loop that generates three plots of `p` versus `se` when the sample sizes equal $N = 25$, $N = 100$, and $N = 1000$.

`@instructions`
- Your for-loop should contain three lines of code to be repeated for three different values of N.
- The first line within the for-loop should use the `seq` function to generate a vector of 100 values of `p` that range from 0 to 1.
- The second line within the for-loop should use the `sqrt` function to generate a vector of standard errors for all values of `p`.
- The last line within the for-loop should use the `plot` function to generate a plot with `p` on the x-axis and `se` on the y-axis.
- Use the `ylim` argument to keep the y-axis limits constant across all three plots. The lower limit should be equal to 0 and the upper limit should equal the highest calculated standard error across all values of `p` and `N`.

`@hint`
- The standard error of the sample average is $$\mbox{SE}(\bar{X}) = \sqrt{p (1-p) / N}$$.
- The standard error of the sample average is highest when p = 0.5 and N is small.
- Use the command `for` to define your loop within parentheses. Write the lines of code to be repeated within curly brackets. For example, the example for-loop executes three lines of code for `i` in the vector `c`.

```{r}
c <- c(1,2,3)

for(i in c){
  example_line1
  example_line2
  example_line3
}
```

`@pre_exercise_code`
```{r}
#no pec
```

`@sample_code`
```{r}
# The vector `p` contains 100 proportions of Democrats ranging from 0 to 1
p <- seq(0, 1, length = 100)

# The vector `sample_sizes` contains the three sample sizes
sample_sizes <- c(25, 100, 1000)

# Write a for-loop that calculates the standard error `se` for every value of `p` for each of the three samples sizes `N` in the vector `sample_sizes`. Plot the three graphs, using the `ylim` argument to standardize the y-axis across all three plots.






```

`@solution`
```{r}
# The vector `p` contains 100 proportions of Democrats ranging from 0 to 1
p <- seq(0, 1, length = 100)

# The vector `sample_sizes` contains the three sample sizes
sample_sizes <- c(25, 100, 1000)

# Write a for-loop that calculates the standard error `se` for every value of `p` for each of the three samples sizes `N` in the vector `sample_sizes`. Plot the three graphs, using the `ylim` argument to standardize the y-axis across all three plots.
for(N in sample_sizes){
  se <- sqrt(p*(1-p)/N)
  plot(p, se, ylim = c(0,0.5/sqrt(25)))
}
```

`@sct`
```{r}
test_correct(
  test_for_loop(cond_test = test_student_typed("N in sample_sizes",
                                               not_typed_msg = "You can use `N in sample_sizes` to define your for loop."),
                expr_test = test_function("sqrt", args = x,
                  not_called_msg = "Have you used `sqrt` to calculate the standard error?",
                  incorrect_msg = "Make sure to specify the correct formula when using the `sqrt` function to calculate standard error."), 
              expr_test = test_function("plot", args = c(x,y,ylim),
                  not_called_msg = "Have you used `plot` to generate a plot `p` and `se`?",
                  args_not_specified = "Remember to specify the vectors that contain the values to be plotted on the x- and y-axes when using the `plot` function.",
                  incorrect_msg = "Make sure to plot `p` on the x-axis and `se` on the y-axis. Specify the limits for the y-axis using `ylim`."),
              not_found_msg = "Make sure to use a for-loop to repeat the standard error calculation and plot for three sample sizes."))
success_msg("Great work! Let's move on to the next exercise.")
```  

## Exercise 7

```yaml
type: MultipleChoiceExercise
lang: r
xp: 100
skills: 1
```

Our estimate for the difference in proportions of Democrats and Republicans is $d = \bar{X} - (1-\bar{X})$. Which derivation correctly uses the rules we learned about sums of random variables and scaled random variables to derive the expected value of $d$?

`@instructions`
$$
\begin{eqnarray}
\mbox{E}[\bar{X} - (1-\bar{X})] &=& \mbox{E}[2\bar{X} - 1] \\
&=& 2\mbox{E}[\bar{X}] - 1  \\
&=& N(2p - 1)\\
&=& Np - N(1-p)
\end{eqnarray}
$$

$$
\begin{eqnarray}
\mbox{E}[\bar{X} - (1-\bar{X})] &=& \mbox{E}[\bar{X} - 1] \\
&=& \mbox{E}[\bar{X}] - 1  \\
&=& p - 1\\
\end{eqnarray}
$$

$$
\begin{eqnarray}
\mbox{E}[\bar{X} - (1-\bar{X})] &=& \mbox{E}[2\bar{X} - 1] \\
&=& 2\mbox{E}[\bar{X}] - 1  \\
&=& 2\sqrt{p(1-p)} -1 \\
&=& p - (1-p)
\end{eqnarray}
$$

$$
\begin{eqnarray}
\mbox{E}[\bar{X} - (1-\bar{X})] &=& \mbox{E}[2\bar{X} - 1] \\
&=& 2\mbox{E}[\bar{X}] - 1  \\
&=& 2p - 1\\
&=& p - (1-p)
\end{eqnarray}
$$
`@hint`
- The expected value of $\bar{X}$ is equal to `p`.

`@sct`
```{r}
msg1 = "Try again. The sample size is not included in this equation."
msg2 = "Try again. Pay attention to the algebra in the first step."
msg3 = "Try again. The expected value of $\bar{X} is equal to `p`."
msg4 = "Nice work!"
test_mc(correct = 4, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

## Exercise 8

```yaml
type: MultipleChoiceExercise
lang: r
xp: 100
skills: 1
```

Our estimate for the difference in proportions of Democrats and Republicans is $d = \bar{X} - (1-\bar{X})$. Which derivation correctly uses the rules we learned about sums of random variables and scaled random variables to derive the standard error of $d$?

`@instructions`

$$
\begin{eqnarray}
\mbox{SE}[\bar{X} - (1-\bar{X})] &=& \mbox{SE}[2\bar{X} - 1] \\
&=& 2\mbox{SE}[\bar{X}]   \\
&=& 2\sqrt{p/N}
\end{eqnarray}
$$

$$
\begin{eqnarray}
\mbox{SE}[\bar{X} - (1-\bar{X})] &=& \mbox{SE}[2\bar{X} - 1] \\
&=& 2\mbox{SE}[\bar{X} - 1]   \\
&=& 2\sqrt{p(1-p)/N} - 1
\end{eqnarray}
$$

$$
\begin{eqnarray}
\mbox{SE}[\bar{X} - (1-\bar{X})] &=& \mbox{SE}[2\bar{X} - 1] \\
&=& 2\mbox{SE}[\bar{X}]   \\
&=& 2\sqrt{p(1-p)/N}
\end{eqnarray}
$$

$$
\begin{eqnarray}
\mbox{SE}[\bar{X} - (1-\bar{X})] &=& \mbox{SE}[\bar{X} - 1] \\
&=& \mbox{SE}[\bar{X}]   \\
&=& \sqrt{p(1-p)/N}
\end{eqnarray}
$$
`@hint`
- The standard error of $\bar{X}$ is equal to $\sqrt{p(1-p)/N}$
- Subtracting a constant does not add any variability, so it does not affect the standard error.

`@sct`
```{r}
msg1 = "Try again. The standard error of $\bar{X}$ is equal to the square root of the spread divided by the sample size."
msg2 = "Try again. Subtracting 1 does not affect the standard error."
msg3 = "Nice work!"
msg4 = "Try again. Pay attention to the algebra on the first line."
test_mc(correct = 3, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

## Exercise 9
Say the actual proportion of Democratic voters is $p=.45$. In this case, the Republican party is winning by a relatively large margin of $d= -0.1$, or a 10% margin of victory. What is the standard error of the spread $2\bar{X}-1$ in this case? 

`@instructions`

`@hint`

`@pre_exercise_code`
```{r}
#no pec
```

`@sample_code`
```{r}
```

`@solution`
```{r}
# `N` represents the number of people polled
N <- 25

# `p` represents the proportion of Democratic voters
p <- 0.45

# Calculate the standard error of the spread. Print this value to the console.
2*sqrt(p*(1-p)/N)


```

`@sct`
```{r}
test_error()
test_function("sqrt", args = "x", eval=NA,
              not_called_msg = "Make sure you are using the `sqrt` function to calculate the standard error.")
test_output_contains("0.1989975", incorrect_msg = "You are not providing a calculation that gives the correct answer. Make sure you are using the formula we derived to calculate the standard error of the spread.")
success_msg("Great work! Let's work on a similar problem.")
```

## Exercise 10
So far we have said that the difference between the proportion of Democratic voters and Republican voters is about 10% and that the standard error of this spread is about 0.2 when $N=25$. Select the statement that explains why this sample size is sufficient or not.

`@instructions`
- This sample size is sufficient because the expected value of our estimate $2\bar{X}-1$ is $d$ so our prediction will be right on.
- This sample size is too small because the standard error is larger than the spread.
- This sample size is sufficient because the standard error of about 0.2 is much smaller than the spread of 10%.
- Without knowing `p`, we have no way of knowing that increasing our sample size would actually improve our standard error.

`@hint`
- A standard error of 0.2 means the proportion of Democratic voters may vary from the estimated value by plus or minus 0.2.

`@sct`
```{r}
msg1 = "Try again. Consider the formula used to calculate standard error. How can standard error be reduced?"
msg2 = "Great work!"
msg3 = "Try again. The spread of 10% means that the proportions differ by 0.1."
msg4 = "Try again. Standard error becomes smaller as you increase the sample size."
test_mc(correct = 3, feedback_msgs = c(msg1,msg2,msg3,msg4))
```

---
## End of Assessment: The Central Limit Theorem

```yaml
type: PureMultipleChoiceExercise
xp: 0
skills: 1
```

This is the end of the programming assignment for this section. No further assessments are accessible from this assignment at this time. Please DO NOT click through to additional assessments from this page. If you do click through, your scores will NOT be recorded.

You can close this window and return to Data Science: Visualization.

*** =instructions


*** =hint
Continue on with the course within the edX platform.