---
title: Parameters and Estimates
title_meta: Chapter 2
description: In this chapter, you will learn about parameters and estimates using the example of election polling.
---
## Exercise 1

Suppose you poll a population in which a proportion $p$ of voters are Democrats and $1-p$ are Republicans. Your sample size is $N=25$. Consider the random variable $S$ which is the **total** number of Democrats in your sample. What is the expected value of this random variable? Hint: it's a function of $p$.

__Answer__
$$25p$$

## Exercise 2
What is the standard error of $S$ ? Hint: it's a function of $p$.

__Answer__
$$ \sqrt{25 p (1-p)}$$

## Exercise 3
Consider the random variable $S/N$. This is equivalent to the sample average, which we have been denoting as $\bar{X}$. What is the expected value of the $\bar{X}$? Hint: it's a function of $p$.

__Answer__

$$p$$


## Exercise 4
What is the standard error of $\bar{X}$? Hint: it's a function of $p$.

__Answer__
$$\sqrt{p (1-p) / 25}$$

## Exercise 5
Write a line of code that gives you the standard error `se` for the problem above for `p <- seq(0, 1, length = 100)`. Make a plot of `se` versus `p`. 

```{r}
p <- seq(0, 1, length = 100)
se <- sqrt(p*(1-p)/25)
plot(p, se)
```

## Exercise 6
Copy the code above and put it inside a for-loop to make the plot for $N=25$, $N=100$ and $N=1000$. 

```{r}
for(N in c(25, 100, 1000)){
  p <- seq(0, 1, length = 100)
  se <- sqrt(p*(1-p)/N)
  plot(p, se, ylim = c(0,0.5/sqrt(25)))
}
```

  

## Exercise 7
If we are interested in the difference in proportions, $p - (1-p)$, our estimate is $d = \bar{X} - (1-\bar{X})$. Use the rules we learned about sums of random variables and scaled random variables to derive the expected value of $d$.

$$
\begin{eqnarray}
\mbox{E}[\bar{X} - (1-\bar{X})] &=& \mbox{E}[2\bar{X} - 1] \\
&=& 2\mbox{E}[\bar{X}] - 1  \\
&=& 2p - 1\\
&=& p - (1-p)
\end{eqnarray}
$$

## Exercise 8
What is the standard error of $d$?

$$
\begin{eqnarray}
\mbox{SE}[\bar{X} - (1-\bar{X})] &=& \mbox{SE}[2\bar{X} - 1] \\
&=& 2\mbox{SE}[\bar{X}]   \\
&=& 2\sqrt{p(1-p)/N}
\end{eqnarray}
$$

## Exercise 9
If the actual is $p=.45$, it means the Republicans are winning by a relatively large margin since $d= -.1$, which is a 10% margin of victory. What is the standard error of $2\hat{X}-1$ in this case? 

```{r}
p <- 0.4
2*sqrt(p*(1-p)/25)
```

## Exercise 10
Given the answer to 9, which of the following best describes your strategy of using a sample size of $N=25$?

  A. The expected value of our estimate $2\bar{X}-1$ is $d$ so our prediction will be right on.
  B. Our standard error is larger than the difference, so the chances of $2\bar{X}-1$ being positive and throwing us off were not that small. We should pick a larger sample size.
  C. The difference is 10% and the standard error is about 0.2, so much smaller than the difference.
  D. Because we don't know $p$, we have no way knowing that making $N$ larger would actually improve our standard error.
